Accelerating adaptive ultrasound algorithms using graphics processing units \vspace{0.5cm}

### 2014-01-19 ThesisJP/Background/background.tex
Finished summary of paper 4 and 5.
Writing discussion.

@@

@@

@@

@@

\subsection{Shift invariance and sampling}

@@

@@

@@

>>>>

Architecture. See Paper II.

@@

Small section about heart anatomy.

Typical probes, scan sequences, resolution and sampling.

@@

Delay and sum. Apodization. (See intro to Paper II or III).
Time v.s. phase delays.

@@

Intro to adaptive beamforming. List other variants (LCA, beamspace, Eigen space etc.).

Beamspace data is typically refers to the polar grid that cardiac ultrasound data is located in prior to scan conversion. In combination with the Capon beamformer, beamspace refers to the K-space representation of the impinging signals (hence the \nom{FFT}{Fast fourier transform} of the channel data).

Not phase aberration correction.

@@

Add section about how to present data (max v.s mean etc.)

\subsection{Shift invariance}

@@

Get section from master theses. Ray casting and opacity functions.

@@

Visibility driven visualization.

@@

Small chapter about different simulation tools (See hos paper).

### 2014-01-19 ThesisJP/Introduction/introduction.tex
Writing discussion.

\section{Discussion}
The main contributions of this thesis are:
\begin{enumerate}
\item a GPU implementation of the Capon beamformer
\item a GPU implementation of the beamspace Capon beamformer
\item that real-time beamspace Capon beamforming is achieved for cardiac ultrasound imaging
\item the first investigation of Capon beamforming applied on multiple frames in medical ultrasound imaging (simulated, \textit{in-vitro}, and \textit{in-vivo})
\item an investigation of shift-invariance when using the Capon beamformer
\item a method for improved shift-invariance of the Capon beamformer
\item a method for reduced blood-pool noise in volume renderings of cardiac ultrasound volumes
\item a GPU implementation of this adaptive volume rendering method
\item a GPU implementation of simple paint-like simulation tool for rapid visualization ultrasound pressure fields
\end{enumerate}

@@

Even with all our effort in \textbf{Paper I} we were only able to achieve a fraction of the maximum throughput of the target GPU. This result shows how hard it is to reach the theoretical level of throughput, and that how close one can get is highly algorithm depended. It might be ways of improving these numbers, but both the solver and covariance estimation step consist of several points where fine granular synchronization and serialized instructions are needed. This will restrict even the most fine tuned implementation from reaching the maximum throughput of the GPU. An interesting observation can be found in Fig.\,10 in \textbf{Paper I}. Here we see how the achieved throughput of the covariance estimation kernel is more than three times higher than the theoretical throughput of a modern, high-end, \nom{CPU}{Central processing unit}. This clearly shows the benefit of modern GPU computing.

Solving was done by a batched Gauss Jordan solver implemented by Nvidia. In our work we focused on the covariance estimation step, and did not try to beat Nvidia at home. Nevertheless, with the measured throughput in mind, a natural next step would be to analyze the current solver, and figure out if any further optimization is possible. Since the sample covariance matrix is hermitian, a solver based on Cholesky decomposition might also help to improve the solver's throughput.
\\\\
In \textbf{Paper II} we present the first medical ultrasound loops processed with ES- and BS-Capon beamforming.  The paper does not provide a detailed evaluation of the method's \textit{in vivo} performance. Yet, initially it has been proven hard to transfer results obtained in simulations to \textit{in vivo} applications. The improvements seen are minimal, especially with the amount of calculation required in mind. The method is clearly able to decrees the lateral with of lateral point scatterers and thin structures in addition to reduce sidelobes cluttering tissue edges. On the other hand, the overall contrast is visually not improved and sometimes worse. Since the Capon beamformer does not reduce the noise floor just directional noise, contrast will get worse in low SNR applications when signal cancellation occurs.

Discuss why cardiac ultrasound is selected and if full Capon will be real-time for other modalities.

Discuss LCA.
\\\\
Paper3
Discuss how contrast might get improved since the amount of signal cancellation is reduced. 
\\\\
Paper4

@@

\section{Conclusion}

Concluding remarks...

\section{Future work}

Merge this section with discussions..?

All five papers...

Detailed investigation of why results obtained in simulations are not directly transferable to \textit{in vivo}.

>>>>

\section{Discussion and future work}
The main contributions of this thesis are:
\begin{enumerate}
\item a GPU implementation of the Capon beamformer
\item a GPU implementation of the beamspace Capon beamformer
\item that real-time beamspace Capon beamforming is achieved for cardiac ultrasound imaging
\item the first investigation of Capon beamforming applied on multiple frames in medical ultrasound imaging (simulated, \textit{in vitro}, and \textit{in vivo})
\item an investigation of shift-invariance when using the Capon beamformer
\item a method for improved shift-invariance of the Capon beamformer
\item a method for reduced blood-pool noise in volume renderings of cardiac ultrasound volumes
\item a GPU implementation of this adaptive volume rendering method
\item a GPU implementation of simple paint-like simulation tool for rapid visualization ultrasound pressure fields
\end{enumerate}

@@

Even with all our effort in \textbf{Paper I} we were only able to achieve a fraction of the maximum throughput of the target GPU. This result shows how hard it is to reach the theoretical level of throughput, and that how close one can get is highly algorithm depended. It might be ways of improving these numbers, but both the solver and covariance estimation step consist of several points where fine granular synchronization and serialized instructions are needed. This will restrict even the most fine tuned implementation from reaching the maximum throughput of the GPU. An interesting observation can be found in Fig.\,10 in \textbf{Paper I}. Here we see how the achieved throughput of the covariance estimation kernel is more than three times higher than the theoretical throughput of a modern, high-end, \nom{CPU}{Central processing unit}. This clearly shows the benefit of modern GPU computing.  Solving was performed by a batched Gauss Jordan solver implemented by Nvidia. In our work we focused on the covariance estimation step, and did not try to beat Nvidia at home. Nevertheless, with the measured throughput in mind, a natural next step would be to analyze the current solver, and figure out if any further optimization is possible. Since the sample covariance matrix is hermitian, a solver based on Cholesky decomposition might also help to improve the solver's throughput.
\\\\
In \textbf{Paper II} we focused on achieving real time Capon processing for cardiac ultrasound, but we also present the first medical ultrasound loops processed with ES- and BS-Capon beamforming.  The paper does not provide any detailed evaluation of the method's \textit{in vivo} performance. Yet, initially is has been proven hard to transfer results obtained in simulations to \textit{in vivo} images. The improvements seen are minimal, especially with the amount of calculations required in mind. The beamformers are clearly able to decrees the lateral width of lateral point scatterers and thin structures, and to sharpen step edges. On the other hand, the overall contrast is visually not improved and sometimes it is even worse. Since the Capon beamformer only affect directional noise, it is clear that contrast will get worse in low \nom{SNR}{Signal-to-noise ratio} applications when signal cancellation occurs. The oversampling approach presented in \textbf{Paper III} might improve on this, but it will only reduce signal cancellation if the phase shifts across the aperture are small and close to linear. Another issue is that bright points with a wide lateral profile exhibits better visual contrast when surrounded by speckle noise than points with narrow lateral profiles. In future work a detailed investigation of why results obtained in simulations are not directly transferable to \textit{in vivo} needs to be performed. It will also be crucial to show larger improvements \textit{in vivio} than what is presented in \textbf{Paper II}. Both to justify all our computations, and to show an image which has improved resolution and contrast. As pointed out at the end of \textbf{Paper II}, cardiac ultrasound might not be the best modality for Capon beamforming. Cardiac ultrasound imaging was selected because it is the medical ultrasound modality which requires most floating points operations per second. An application where detection of closely spaced point targets are of importance might be better suited. A better suited adaptive beamformer for cardiac applications is the \nom{LCA}{Low complexity beamformer} beamformer, where the Capon optimization problem is applied on a set of predefined windows. This method has linear complexity and shares many properties with the Capon beamformer. 
\\\\
Paper3
Discuss how contrast might get improved since the amount of signal cancellation is reduced.

Oversampling on receive will increase the penalty of processing beamspace data. Scan conversion will also take more time, and the end result should probably have higher resolution to provide the sharp definitions provided by Capon beamforming. 
\\\\
Paper4

@@
